Voting Patterns and Socio-Economic: A Twenty-Year Analysis of Brazil's Developmental Dynamics

### 2024-02-15 outputs/paper/paper.qmd
added knitr

---
title: "Voting Patterns and Socio-Economic Tides: A Twenty-Year Analysis of Brazil's Developmental Dynamics"
author: "Hao Wang / Siqi Fei / Xuecheng Gao"
thanks: "Code and data are available at: https://github.com/groundUofT888/Term-Paper-2.git ; Replication on Social Science Reproduction platform available at: www.aeaweb.org/articles?id=10.1257/app.20210529"
date: "February 12, 2024"
date-format: long
format: 
  pdf: 
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    documentclass: article
    geometry: margin = 1in
    abstract: " "
    number-sections: true
output: 
  bookdown::pdf_document2
toc: false
fig_caption: yes 
nocite: '@*'
bibliography: references.bib
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("readr")
# install.packages("here")
# install.packages("dplyr")
# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("knitr")

library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)

data_path <- here("outputs", "edited_data", "election_cleaned.csv")
election <- read.csv(data_path)
data_path2 <- here("outputs", "edited_data", "schools.csv")
school <- read.csv(data_path2)
```
\newpage
\begin{center}
\textbf{\large Abstract}
\end{center}

\begin{quote}
This paper presents a comprehensive analysis of socio-economic data for Brazil spanning from 1996 to 2016. The data, representing over 33,000 Brazilians across six electoral cycles, encompasses economic indicators such as income per capita, low-income rates, and unemployment rates, as well as political variables like electoral years and district codes. Through meticulous data visualization and statistical analysis, we examine the distributions of these key indicators and their interplay with electoral processes.

Our findings reveal diverse economic conditions as indicated by the spread of income per capita and a bimodal distribution in low-income rates, suggesting the existence of two distinct economic subgroups within the population. Furthermore, the unemployment rate exhibits a concentration around lower rates with a tail extending towards higher values, highlighting the prevailing low unemployment with instances of higher rates.

By integrating these economic measures with educational data like dropout rates, we uncover complex correlations and the varying impact of voter numbers on educational outcomes. We discuss the implications of these relationships, exploring how electoral districts and voter numbers correlate with socio-economic conditions.

This study provides valuable insights into the socio-economic and political dynamics of Brazil, emphasizing the importance of electoral systems in shaping economic and educational landscapes. It contributes to a deeper understanding of how policy and governance can influence socio-economic development and underscores the need for nuanced policy interventions to address economic disparities and promote equitable growth.
\end{quote}

\newpage
# Introduction
## Two-round elections in Brazil
Brazil's two-round electoral system is a relatively common electoral method. The mode of election is to select the winner after two votings. In the first-round voters can choose the candidate they support among multiple candidates. If one of the candidates gets more than half of the votes in the first round, then he can be directly elected. If no candidate receives more than half of the votes in the first round, a second round will take place. In the second round of voting, voters will choose between the two candidates who received the highest votes in the first round.
Election regulations：1. Voters must be registered at least 151 days before the election. 2. Voting is compulsory for all literate Pakistani citizens between the ages of 18 and 69. 3. The election is held on the first Sunday in October.

## Economic and social issues in Brazil
Major social and economic issues in Brazil. First, Brazil’s poverty gap. There are many depressed areas in Brazil. Many people live below the poverty line and lack basic living conditions such as housing, education, medical care, and jobs. This leads to social instability. Second, the high crime rate. Some cities in Brazil have high crime rates and violence problems, such as robberies, gun proliferation, and drug trafficking. These problems mainly stem from poverty caused by uneven distribution of wealth and the deterioration of local security. Third, Brazil has high unemployment. The number of unemployed people is increasing in Brazil, and many structural unemployment results in high unemployment rates for a large number of low-skilled laborers. Also lower wages for many people worsen poverty and inequality. Fourth, there are issues with the education system. Brazil's education system has many problems, such as uneven education quality, uneven allocation of government funds for education, shortage of professors, and high dropout rates.

## Overview
In 2022, Brazil's presidential election broke new grounds with an unprecedented turnout of over 150 million voters, capturing the attention of national and international media. The world had seen not just about the political tactics of the candidates, but also the public's vested interest in how these elections could shape their day-to-day lives. This paper aims to study the influence of Brazil's electoral system on three critical socio-economic indicators: income level, unemployment rate and educational resources.With a comprehensive look at recent data and trends, we will investigate how the details of voting and representation create a chain reaction throughout the economy, affecting everything from job opportunities to the quality of education. By collecting together statistical analysis with real-world implications, we'll provide a clearer picture of the electoral system's impact on the Brazilian populace and discuss the broader implications of these findings on social equality and economic stability.

@@

It uses data sets constructed from four data sources: - 1) Brazil municipal election data (Tribunal Superior Eleitoral, 1996-2016); - 2) Brazil demographic censuses (Institute Brasileiro de Geografia e Estatística, 1980, 1991, 2000, 2010); - ) Brazil school census (Ministério da Educação, 1997-2016); - 4) nighttime lights (National Oceanic and Atmospheric Administration, 1997-2016a,-)

## Methodology

@@

**election.csv/election_clean.csv**

The data set offers a comprehensive longitudinal collection of socio-economic indicators for Brazil from 1996 to 2016. Over the course of these twenty years, the data set captures the socio-economic fluctuations and trends across six Brazilian electoral cycles, with a granular focus on a variety of indicators.

-   **Time frame**: Spanning two decades, the data set provides insights into the socio-economic changes over time, coinciding with multiple electoral events in Brazil.

-   **Volume**: Each election year is detailed with 5,567 data entries, resulting in a total of 33,188 entries over the entire period, allowing for in-depth analysis and trend identification.

-   **Indicators Included**:

    -   `cut_dist`: A geographical identifier for districts or electoral cuts, offering spatial analysis opportunities.
    -   `elect_year`: The year of the election, anchoring the socio-economic data within a political timeline.
    -   `inc_post`: Per capita income figures post-election, reflecting the general economic well-being.
    -   `inc_0_50_post`: Income figures for the bottom 50% post-election, highlighting the economic status of the less affluent demographic.
    -   `edu_univ`: A variable intended to capture educational attainment or resources, yet unspecified in the initial data extract.
    -   `unempl_post`: Post-election unemployment rates, indicating labor market conditions.
    -   `tse_code`: Potentially linked to electoral zones or districts, allowing for analysis of electoral influence on socio-economic aspects.

**schools.csv**

The dataset includes an array of variables that shed light on the multifaceted nature of Brazil's socio-economic and political landscape:

- **Electoral Dynamics**: With variables like `elect_year` and `tse_code`, the dataset chronicles the electoral events, reflecting the evolving political scene at regular four-year intervals.

- **Geographic and Demographic Insights**: Indicators such as `ibge7_code`, `cut_dist`, and `pop_den_pre` provide detailed geographic and demographic information, highlighting regional disparities and population distribution.

- **Economic Measures**: Economic variables such as `dep_pca_ptile_mean`, `eq_pca_ptile_mean`, `inc_T0`, and `inc_0_50_T0` reveal insights into the income levels, economic stratification, and wealth distribution within Brazil.

- **Educational Outcomes**: Metrics like `MAT_DROP_mean`, `MAT_PASS_mean`, and `MAT_FAIL_mean` offer a glimpse into the educational sector's performance, a critical aspect of Brazil's socio-economic fabric.

- **Social Indicators**: Social factors such as illiteracy rates (`illit_T0`), unemployment figures (`unempl_T0`), and measures of inequality (`gini_T0`) provide a deeper understanding of the societal challenges that have shaped Brazil's development over the years.

\renewcommand{\arraystretch}{2}

\newpage
# Result

Our statistical analysis began with a computation of the sample size which represents the population of Brazilians in 2022 election, with a population size of 203.1 million. We determined the sample size as 33188 in this study, ensuring the analysis result would be within an acceptable margin of error. This sample size was substantiated by Quota sampling, ensuring the representations of our data set.

The sample we used were randomly selected every 4 year from 1996 to 2016. Each year, there are 5567 Brazilians were chose to be the sample set. we generated all the sample sets and run the data analysis to study the factors.

@@

## Overview
```{r}
#| label: fig-hist
#| fig-cap: histogram
#| echo: false
#| message: false
#| warning: false
#| fig.width: 8
#| fig.height: 7

@@

In the paper, Moya Chin uses the regression discontinuity design to compare single-round and two-round elections in Brazil. Our paper makes two findings based on Moya Chin's data and theory.

1. The two rounds of elections in Brazil have a good impact on Brazil's downstream economy.

2. Candidates prefer to attract low-income people and people with low education during the election.

We found from Moya Chin's paper that politicians' behavior during elections affects the provision of local public facilities, but the allocation of these resources changes when politicians are elected.

We find that two-round elections can incentivize candidates to provide resources for downstream economic and public education. In terms of education, the dropout rate has been significantly reduced and the literacy rate has increased. According to census data in 2000 and 2010, the primary school literacy rate has been on an upward trend, while the dropout rate has been on a downward trend. According to the school census, the school's infrastructure has been significantly improved during the campaign, and public facilities such as the school's library, gymnasium, and classrooms have received more funding. Resource levels in public schools have increased and differences between schools have become smaller.

In terms of the provision of public goods, we found that in the two-round election of urban school resource allocation compared with the single-round municipal government, these schools receive more resources in terms of equipment and infrastructure, and these resources are more distributed among schools. fair. Politicians provide a large number of products such as computers, photocopiers, projectors, etc. These products reduce dropout rates and increase literacy rates. These grants have improved local education standards and have a positive impact on local education. This shows that two rounds of elections will bring about more public goods provision, which will have a positive impact on education. These improvements in public education are important for long-term socioeconomic development because better education can create a more highly skilled workforce in the future and may reduce structural unemployment in the future.

However, we find that two rounds of elections have limited economic impact in the mid-term. According to the Figure, it can be seen that per capita income increased during the election period, and the number of low-income people decreased compared with before. But these results were not evident in the statistics. This suggests that while two rounds of elections will improve the downstream economy, changes in income levels and unemployment will take longer to materialize.

We also found politicians in Brazil using political manipulation and tactics in elections. In Brazil, low-income groups make up a larger share of the population, so attracting votes from this group could have a significant impact on election results. Therefore, even if high-income groups pay more attention to elections than low-income groups, politicians will allocate more resources to the majority of voters.

We believe that Brazil’s two rounds of elections will bring short-term benefits to the downstream economy and education. However, long-term changes in Brazil’s economic and social problems require long-term observation.

>>>>

---
title: "Voting Patterns and Socio-Economic: A Twenty-Year Analysis of Brazil's Developmental Dynamics"
author: "Hao Wang / Siqi Fei / Xuecheng Gao"
thanks: "Code and data are available at: https://github.com/groundUofT888/Term-Paper-2.git ; Replication on Social Science Reproduction platform available at: https://doi.org/10.48152/ssrp-qxvj-fz21"
date: "February 12, 2024"
date-format: long
format: 
  pdf: 
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    documentclass: article
    geometry: margin = 1in
    abstract: " "
    number-sections: true
output: 
  bookdown::pdf_document2
toc: false
fig_caption: yes 
nocite: '@*'
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---
\newpage
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("readr")
# install.packages("here")
# install.packages("dplyr")
# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("knitr")

library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(knitr)

data_path <- here("outputs", "edited_data", "election_cleaned.csv")
election <- read.csv(data_path)
data_path2 <- here("outputs", "edited_data", "schools.csv")
school <- read.csv(data_path2)
```

```{=tex}
\begin{center}
\textbf{\large Abstract}
\end{center}
```
```{=tex}
\begin{quote}
The paper replicates data to analyze economic indicators such as income per capita, the low-income rate, and the unemployment rate during two-round elections in Brazil, while considering the dropout rate as a main educational indicator and exploring the impact of electoral provisions on citizens, the country's economy, and society. We conduct secondary research on Brazil’s economic and education data and find that politicians’ campaign strategies influence voters in two-round electoral systems, further affecting the local economy and policy implementation.

We hope that by analyzing the Brazilian electoral system, we can provide new perspectives and ideas for studying the impact of the world's political electoral systems. We call for future investigations in this area to provide more information, and we also look forward to offering useful suggestions and inspirations for Brazil's social development and economic construction.
\end{quote}
```

# Introduction

In 2022, Brazil's presidential election broke new grounds with an
unprecedented turnout of over 150 million voters, capturing the
attention of national and international media. The world had seen not
just about the political tactics of the candidates, but also the
public's vested interest in how these elections could shape their
day-to-day lives. This paper aims to study the influence of Brazil's
electoral system on three critical socio-economic indicators: income
level, unemployment rate and educational resources.With a comprehensive
look at recent data and trends, we will investigate how the details of
voting and representation create a chain reaction throughout the
economy, affecting everything from job opportunities to the quality of
education. By collecting together statistical analysis with real-world
implications, we'll provide a clearer picture of the electoral system's
impact on the Brazilian populace and discuss the broader implications of
these findings on social equality and economic stability.

# Background

## Two-round elections in Brazil

Brazil's two-round electoral system is a relatively common electoral
method. The mode of election is to select the winner after two votings.
In the first-round voters can choose the candidate they support among
multiple candidates. If one of the candidates gets more than half of the
votes in the first round, then he can be directly elected. If no
candidate receives more than half of the votes in the first round, a
second round will take place. In the second round of voting, voters will
choose between the two candidates who received the highest votes in the
first round. Election regulations：1. Voters must be registered at least
151 days before the election. 2. Voting is compulsory for all literate
Pakistani citizens between the ages of 18 and 69. 3. The election is
held on the first Sunday in October.

## Economic and social issues in Brazil

Major social and economic issues in Brazil. First, Brazil’s poverty gap.
There are many depressed areas in Brazil. Many people live below the
poverty line and lack basic living conditions such as housing,
education, medical care, and jobs. This leads to social instability.
Second, the high crime rate. Some cities in Brazil have high crime rates
and violence problems, such as robberies, gun proliferation, and drug
trafficking. These problems mainly stem from poverty caused by uneven
distribution of wealth and the deterioration of local security. Third,
Brazil has high unemployment. The number of unemployed people is
increasing in Brazil, and many structural unemployment results in high
unemployment rates for a large number of low-skilled laborers. Also
lower wages for many people worsen poverty and inequality. Fourth, there
are issues with the education system. Brazil's education system has many
problems, such as uneven education quality, uneven allocation of
government funds for education, shortage of professors, and high dropout
rates.

@@

It uses data sets constructed from four data sources: - 1) Brazil
municipal election data (Tribunal Superior Eleitoral, 1996-2016); - 2)
Brazil demographic censuses (Institute Brasileiro de Geografia e
Estatística, 1980, 1991, 2000, 2010); - 3) Brazil school census
(Ministério da Educação, 1997-2016); - 4) nighttime lights (National
Oceanic and Atmospheric Administration, 1997-2016a,-)

@@

### Data Sets Overview

Our comprehensive data collection provides a longitudinal perspective on
the socio-economic and political landscape of Brazil, spanning from 1996
to 2016. This period covers six Brazilian electoral cycles, offering a
unique vantage point for analyzing socio-economic fluctuations and
trends through detailed indicators.

### 1. Election Dataset (election_cleaned.csv)

-   Path: outputs/edited_data/election_cleaned.csv

-   Description: Contains socio-economic indicators for Brazil,
    capturing fluctuations and trends across six electoral cycles from
    1996 to 2016.
    ```{r echo = FALSE, results = 'asis'}
    election_vars <- tibble(
      Variable = c("cut_dist", "elect_year", "inc_post", "inc_0_50_post", "edu_univ", "unempl_post", "tse_code"),
      Description = c("Geographical identifier for districts.",
                      "Year of the election.",
                      "Per capita income post-election.",
                      "Income for the bottom 50% post-election.",
                      "Educational attainment/resources.",
                      "Post-election unemployment rates.",
                      "Electoral zones/districts identifier.")
    )

    kable(election_vars, format = "markdown", caption = "Election Dataset Variables")

    ```
\newpage
    ### 2. Schools Dataset (schools.csv)

-   Path: outputs/edited_data/schools.csv

-   Description: Offers insights into Brazil's socio-economic and political landscape through educational and economic measures.
```{r echo = FALSE, results ='asis'}

# Schools Dataset Variables
schools_vars <- tibble(
  Variable = c("ibge7_code", "cut_dist", "pop_den_pre", "dep_pca_ptile_mean", "MAT_DROP_mean"),
  Description = c("Geographic identifier (7-digit IBGE code).",
                  "Geographical identifier for districts.",
                  "Pre-election population density.",
                  "Mean dependency ratio across percentiles.",
                  "Mean drop-out rate in schools.")
)

kable(schools_vars, format = "markdown", caption = "Schools Dataset Variables")

```
\renewcommand{\arraystretch}{2}

# Result

The sample size, comprising 33,188 observations, was determined to ensure representativeness and statistical validity. The dataset encompasses 5,567 data points per election year, reflecting a robust cross-section of Brazil's population and socio-economic conditions during the study period.

@@

## Overview

```{r}
#| label: fig-hist
#| fig-cap: "overview histogram"
#| echo: false
#| message: false
#| warning: false
#| fig.width: 8
#| fig.height: 7

@@

Our examination of the data presented by Moya Chin through a regression discontinuity design yields nuanced insights into the effects of Brazil's electoral system on socio-economic variables. We delineate two primary findings from our analysis:

* Electoral Influence on Economic Conditions: Contrary to the initial reading of the data, the two-round elections in Brazil do not show a definitive positive impact on the broader downstream economy within the period studied. While there is an observable increase in per capita income and a decrease in the number of low-income individuals during election periods, these trends are not statistically significant when considering the larger socio-economic context. The data suggests that any changes in income levels and unemployment rates associated with the electoral process are subtle and may require a more extended period to be fully realized and understood.

* Campaign Strategies and Voter Demographics: Our findings align with the hypothesis that candidates during elections strategically target demographics that are most likely to influence the outcome, specifically low-income individuals and those with lower educational attainment. This is reflected in the allocation of resources, such as improvements in school infrastructure and public education facilities, which are more likely to directly benefit these groups. While this approach does translate into immediate educational gains, such as reduced dropout rates and increased literacy rates, it is less clear how these short-term improvements translate into long-term socio-economic development.

From Moya Chin's work, it is evident that the behavior of politicians during election campaigns has a tangible impact on the distribution of public resources, which can shift dramatically post-election. While two-round elections may encourage candidates to commit more resources to public education and economic support, this study finds that the long-term economic effects, such as sustained increases in income or reductions in unemployment, are not as immediate or discernible within the data.

Additionally, the presence of political tactics aimed at low-income voters suggests a complex interplay between electoral strategies and voter demographics. Despite the potential short-term benefits to public education and economic assistance that may arise from two-round elections, our analysis indicates that addressing Brazil's economic and social challenges will likely require a more extended and multifaceted approach.

As the conclusion, while the electoral system does have some influence on socio-economic outcomes, the extent and durability of these effects are contingent on a variety of factors and warrant further, longitudinal study to fully comprehend their scope and longevity.
